Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-10-23 causal-bandit/nips_2016/additional_experiments.tex
updates plot bounds and figures

In this section we compare performance of our algorithms with the standard bandit algorithms on the confounded graph in figure \ref{fig:parallel_confounded}. All the variables are binary and the action space consists of the set of single variable interventions plus the do nothing action, $\calA = \set{\set{do(X_i = j)} \cup \set{do(Z = j)} \cup \set{do()}: 1\leq i \leq N,\; j \in \set{0,1}}$. We choose this setting as it can be seen as a generalisation of the parallel bandit case and allows for factorizations that enable us to compute the exact reward distribution and interventional distributions for large $N$ (in general inference in graphical models is exponential in $N$). As before, we show the average regret over 10,000 simulations with error bars showing three standard errors.

@@

As in the parallel bandit case, we let $Y$ depend only on $X_1$, $P(Y|do(X_1=1)) = \frac{1}{2} + \epsilon$ amd $P(Y|do(X_1=0)) = \frac{1}{2}-\epsilon'$, where $\epsilon' = \epsilon\frac{P(X_1=1)}{P(X_1=0)}$. We vary $N_1$ from 2 to $N$ to vary $m$. The values for the CPD's have been chosen such that the reward distribution is independent of $m$ and so that we can analytically calculate $\eta*$. This allows us to just show the dependence on $m$, removing the noise associated with different models selecting values for $\eta*$ with the same $m$ (and thus same worst case performance) but different performance for a given reward distribution.

@@

In figure \ref{fig:simple_vs_T_misspecified} we exclude actions on $Z$ from the set of allowable actions to demonstrate how algorithm 1 can fail in the presence of a confounding variable. Algorithm 1 incorrectly assumes $P(Y|do(X)) = P(Y|X)$ leading to biased estimates. We let $P(Z) = .6$, $P(Y|\boldsymbol{X}) = X_7 \oplus X_N$ and $P(X|Z)$ be given by:

@@

\begin{figure}[H]
    \begin{subfigure}[t]{0.3\textwidth}
		\centering    
    		\includegraphics[width=\textwidth]{experiment4_20161020_1254.pdf}
    		\caption{Simple regret vs $m(\eta*)$ for fixed horizon $T=400$ and number of variables $N = 50$}
        \label{fig:simple_vs_m_general}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.3\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{experiment7_20161020_1257.pdf}
    		\caption{Simple regret vs horizon, $T$, with $N = 50$ and $m(\eta*)=3.1$ }
        \label{fig:simple_vs_T_general}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.3\textwidth}
    		\centering
    		\includegraphics[width=\textwidth]{experiment5_20161021_2026.pdf}
    		\caption{Simple regret vs horizon, $T$, with $N = 21$, $m=4.3$ with no actions setting $Z$}
    		\label{fig:simple_vs_T_misspecified}
    \end{subfigure}
    \caption{Experimental results on the confounded graph}
    \label{fig:experiments_confounded}
\end{figure}

>>>>

In this section we compare performance of the general algorithm with the standard bandit algorithms on the confounded graph in figure \ref{fig:parallel_confounded}. All the variables are binary and the action space consists of the set of single variable interventions plus the do nothing action, $\calA = \set{\set{do(X_i = j)} \cup \set{do(Z = j)} \cup \set{do()}: 1\leq i \leq N,\; j \in \set{0,1}}$. We choose this setting as it can be seen as a generalisation of the parallel bandit case and allows for factorizations that enable us to compute the exact reward distribution and interventional distributions for large $N$ (in general inference in graphical models is exponential in $N$). As before, we show the average regret over 10,000 simulations with error bars showing three standard errors.

@@

As in the parallel bandit case, we let $Y$ depend only on $X_1$, $P(Y|do(X_1=1)) = \frac{1}{2} + \epsilon$ and $P(Y|do(X_1=0)) = \frac{1}{2}-\epsilon'$, where $\epsilon' = \epsilon\frac{P(X_1=1)}{P(X_1=0)}$. We vary $N_1$ from 2 to $N$ to vary $m$. The values for the CPD's have been chosen such that the reward distribution is independent of $m$ and so that we can analytically calculate $\eta*$. This allows us to just show the dependence on $m$, removing the noise associated with different models selecting values for $\eta*$ with the same $m$ (and thus same worst case performance) but different performance for a given reward distribution.

@@

In figure \ref{fig:simple_vs_T_misspecified} we additionally show the performance of Algorithm 1. We exclude actions on $Z$ from the set of allowable actions to demonstrate how Algorithm 1 can fail in the presence of a confounding variable, as it incorrectly assumes $P(Y|do(X)) = P(Y|X)$ leading to biased estimates. We let $P(Z) = .6$, $P(Y|\boldsymbol{X}) = X_7 \oplus X_N$ and $P(X|Z)$ be given by:

@@

\begin{figure}[H]
    \begin{subfigure}[t]{0.3\textwidth}
		\centering    
    		\includegraphics[width=\textwidth]{experiment4_20161023_2120.pdf}
    		\caption{Simple regret vs $m(\eta*)$ for fixed horizon $T=400$ and number of variables $N = 50$}
        \label{fig:simple_vs_m_general}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.3\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{experiment7_20161020_1257.pdf}
    		\caption{Simple regret vs horizon, $T$, with $N = 50$ and $m(\eta*)=3.1$ }
        \label{fig:simple_vs_T_general}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.3\textwidth}
    		\centering
    		\includegraphics[width=\textwidth]{experiment5_20161023_2118.pdf}
    		\caption{Simple regret vs horizon, $T$, with $N = 21$, $m(\eta*)=4.3$ with no actions setting $Z$}
    		\label{fig:simple_vs_T_misspecified}
    \end{subfigure}
    \caption{Experimental results on the confounded graph}
    \label{fig:experiments_confounded}
\end{figure}
